---
title: 'Homework #3'
author: "Sneha Mehta"
date: "2022-10-15"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
```

## Problem 1




## Problem 2
```{r prob 2 part 1}
accel_data = read_csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(day_type = ifelse(day == "Saturday" | day == "Sunday", "weekend", "weekday")) %>% 
  relocate(day_type, .after = day) %>% 
  pivot_longer(activity_1:activity_1440,
               names_to = "minute",
               names_prefix = "activity_",
               values_to = "activity_count") %>% 
  mutate(minute = as.numeric(minute))
```
This dataset describes five weeks of accelerometer data collected from a 63 year old male with BMI 25. Activity counts were collected each minute of the days. The tidied dataset has 50,400 observations across six variables: the week and day of data collection (day is listed numerically, by day of the week, and classified as a weekday or weekend), the minute of the day the activity count was collected, and the activity count.

```{r prob 2 part 2}
accel_data_sum = accel_data %>% 
  group_by(week, day) %>% 
  summarise(day_activity_total = sum(activity_count)) %>% 
  pivot_wider(
    names_from = day,
    values_from = day_activity_total
  ) %>% select("week", "Monday","Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

accel_data_sum %>% group_by() %>% 
  summarise(across(Monday:Sunday, mean)) %>% view()
```

On average, Saturdays had the lowest activity counts for the individual. Wednesdays - Fridays had the highest.
