---
title: 'Homework #3'
author: "Sneha Mehta"
date: "2022-10-15"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
library(ggridges)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1
```{r prob 1}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

This dataset includes 1,384,617 observations across 15 variables. It describes orders placed on Instacart by 131,209 unique customers. Variables in the dataset include the time each order was placed (day of the week and hour of the day), names of products in the order, isle and department products are found (number and name (i.e., soft drinks, dairy, produce, household goods)), and days since last order.

```{r prob 1 part 1}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n)) %>% view()
```
There are 134 aisles. The aisles with the most number of items ordered are fresh vegetables and fresh fruits at 150609 and 150473 items respectively.

```{r prob 1 part 2}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```
This plot shows all the aisles with at least 10,000 items bought from them. Most aisles have 40,000 items or less purchased from them.

```{r prob 1 part 3}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```
The most popular items from the packaged vegetables & fruits, baking ingredients, and dog food care aisles are organic baby spinach (purchased 9784 times), light brown sugar (499 times), and Snack Sticks Chicken & Rice Recipe Dog Treats (30 times) respectively.

```{r prob 1 part 4}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```
Generally speaking, Pink Lady Apples are purchased earlier in the day than Coffee Ice Cream except for on Day 5

## Problem 2
```{r prob 2 part 1}
accel_data = read_csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(day_type = ifelse(day == "Saturday" | day == "Sunday", "weekend", "weekday")) %>% 
  relocate(day_type, .after = day) %>% 
  pivot_longer(activity_1:activity_1440,
               names_to = "minute",
               names_prefix = "activity_",
               values_to = "activity_count") %>% 
  mutate(minute = as.numeric(minute))
```
This dataset describes five weeks of accelerometer data collected from a 63 year old male with BMI 25. Activity counts were collected each minute of the days. The tidied dataset has 50,400 observations across six variables: the week and day of data collection (day is listed numerically, by day of the week, and classified as a weekday or weekend), the minute of the day the activity count was collected, and the activity count.

```{r prob 2 part 2}
accel_data_sum = accel_data %>% 
  group_by(week, day) %>% 
  summarise(day_activity_total = sum(activity_count)) %>% 
  pivot_wider(
    names_from = day,
    values_from = day_activity_total) %>% 
  select("week", "Monday","Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

accel_data_sum %>% group_by() %>% 
  summarise(across(Monday:Sunday, mean)) %>% view()
```

On average, Saturdays had the lowest activity counts for the individual. Wednesdays - Fridays had the highest.

```{r prob 2 part 3}
ggplot(accel_data, aes(x = minute/60, y = activity_count)) + geom_line(aes(color = day)) + geom_point(aes(color = day), alpha = 0.5) + geom_smooth() + labs(
  title = "Daily Activity Levels",
  x = "Hour",
  y = "Activity Count"
)
```
From this plot, we can see that most activity for this person took place between ~8:00 - ~21:00 hours each day. Activity dipped quite sharply at around 21:00 hours each day and gradually rose from ~5:00 to ~9:00 hours and mostly plateaued for the middle part of each day.

## Problem 3

```{r prob 3 part 0}
library(p8105.datasets)
data("ny_noaa") 
```
This dataset has 2,595,176 observations across 7 variables. It contains data from all New York state weather stations across 5 metrics: precipitation, snowfall, snow depth, maximum temperature, and minimum temperature. It also includes date of observation and the id number of the weather station. There are large amounts of missing observations across all 5 of the measurement metrics.

```{r prob 3 part 1}
ny_noaa = ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), sep = "-", convert = TRUE) %>% 
  mutate(tmax = as.numeric(tmax)/10, 
         tmin = as.numeric(tmin)/10,
         prcp = prcp/10)

table(ny_noaa$snow)
sort(table(ny_noaa$snow), decreasing = TRUE) %>% view()
```
The most commonly observed value for snowfall is 0mm due to the fact that the snowy season is mainly concentrated in a few months (Dec-Feb) (and even then it doesn't snow every day).


```{r prob 3 part 2}
ny_noaa %>% 
  filter(month == 1 | month == 7) %>% mutate(month = case_when(month == 1 ~ "January",
                                                               month == 7 ~ "July")) %>% 
  group_by(id, year, month) %>% 
  mutate(avgtmax = mean(tmax)) %>% 
  ggplot(aes(x = year, y = avgtmax)) + geom_line(aes(color = id), alpha = 0.5) + facet_grid(.~ month) + theme(legend.position = "none") + labs(
  title = "Average Max. Temperature by Month per Station",
  x = "Time",
  y = "Average Max. Temp"
)
```
From the plots, we can see that the average temperature in January is ~27 degrees lower than the average temperature in July. The temperatures in January generally range between around -10 to 10 degrees Celsius, and in July, the temperatures generally range from 20-35 degrees Celcius. In July, we can see one obvious outlier where in ~1988 one station recorded temperatures around 15 degrees.

```{r prob 3 part 3}
min_max = ggplot(ny_noaa, aes(x = tmin, y = tmax)) + geom_hex() + labs(
  x = "Minimum Temp",
  y = "Maximum Temp", 
  title = "Maximum Temp vs Minimum Temp in New York 1981 - 2010"
)
min_max

snowfall = ny_noaa %>% 
  filter(snow %in% (1:99)) %>% 
  ggplot(aes(x = snow, y = as.factor(year), group = year)) + geom_density_ridges() + labs(
    x = "Snowfall(mm)",
    y = "Year",
    title = "Snowfall by Year in New York 1981 - 2010"
  )
snowfall

min_max + snowfall
```

Within the time range, the highest number of days had a minimum temperature between roughly 0-15 degrees Celsius and a maximum temperature between roughly 0-30 degrees. Across the time range, the most common level of snowfall measured per day was between 0-30 mm. There was also a peak around 40-50mm. 
